---
title: "Titanic_Survival"
author: "Louis_Yansaud"
date: "January 29, 2018"
categories: ["R"]
tags: ["R Markdown", "Supervised Machine Learning", "Data Analysis"]
---


##### Loading Data set and libraries
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(caret)
library(Hmisc)
library(randomForest)

train <- read.csv("train.csv")
test <- read.csv("test.csv")
```


##### Data exploration
```{r, message=FALSE, warning=FALSE}
summary(test)
summary(train)
#We can see that survived is missing on the test data, let's input it as a NA value for now

test <- test %>%
  mutate(Survived = NA)

#Lets merge the two dataset together for data wrangling/cleaning
merge_data <- rbind(train, test)

#Checking if there are no duplicates
length(unique(merge_data$PassengerId)) == 891 + 418
#TRUE


```

Looking at the Survived by Sex
```{r, message=FALSE, warning=FALSE}
train %>%
  ggplot(aes(x = Sex, fill = as.factor(Survived)))+
  geom_bar()+
  geom_text(stat = 'count', aes(label = ..count..))+
  scale_fill_discrete(name = 'Survived')
```
Looking at the Survived by Pclass
```{r, message=FALSE, warning=FALSE}
train %>%
  ggplot(aes(x = as.factor(Pclass), fill = as.factor(Survived)))+
  geom_bar()+
  scale_fill_discrete(name = 'Survived')+
  xlab("Pclass")
```

LOoking at the Survived by Sibsp
```{r, message=FALSE, warning=FALSE}
train %>%
  ggplot(aes(x = as.factor(SibSp), fill = as.factor(Survived)))+
  geom_bar()+
  scale_fill_discrete(name = 'Survived')+
  xlab("SibSp")
```

Looking at the Survived by Sibsp and Sex
```{r, message=FALSE, warning=FALSE}
train %>%
  ggplot(aes(x = as.factor(SibSp), fill = as.factor(Survived)))+
  geom_bar()+
  facet_wrap(~Sex)+
  scale_fill_discrete(name = 'Survived')+
  xlab("SibSp")
```


Looking at the Survived by Pclass and Sex
```{r, message=FALSE, warning=FALSE}
train %>%
  ggplot(aes(x = as.factor(Pclass), fill = as.factor(Survived)))+
  geom_bar()+
  facet_wrap(~Sex)+
  scale_fill_discrete(name = 'Survived')+
  xlab("Pclass")
```

Looking at the age distribution
```{r, message=FALSE, warning=FALSE}
merge_data %>%
  ggplot(aes(x = Age))+
  geom_histogram(fill = "#000099", color = "black")
```

###### Data understanding and Data cleaning
Lets group the age into bucket
```{r, message=FALSE, warning=FALSE}


#Lets convert some of the columns name into factors

merge_data$Sex <- as.factor(merge_data$Sex)
merge_data$Pclass <- as.factor(merge_data$Pclass)

#First lets replace the missing values of age by the mean of the subgroup by pclass since it has no missing values 

merge_data %>%
  filter(!is.na(Age)) %>%
  group_by(Pclass) %>%
  summarise(Age_Pclass_mean = round(mean(Age)))

# 1                 39
# 2                 30
# 3                 25

#now lets replace the missing age values

merge_data_missing_age <- merge_data %>%
                            filter(is.na(Age)) %>%
                            mutate(Age = ifelse(Pclass == "1",  39 ,
                                                ifelse(Pclass == "2", 30, 25))) %>% 
                            select(PassengerId, Age)

merge_data <- merge_data %>%
  left_join(merge_data_missing_age, by = 'PassengerId') %>%
  mutate(Age = ifelse(is.na(Age.x), Age.y, Age.x)) %>%
  select(-Age.x, -Age.y)

merge_data$Age_bucket = ifelse(merge_data$Age < 12 , "<12",
                               ifelse(merge_data$Age >= 12 & merge_data$Age < 21, "16-21",
                                ifelse(merge_data$Age >= 21 & merge_data$Age < 30, "21-30",
                                       ifelse(merge_data$Age >= 30 & merge_data$Age < 40, "30-40",
                                              ifelse(merge_data$Age >= 40 & merge_data$Age < 50, "40-50","50+")))))

merge_data %>%
  group_by(Age_bucket) %>%
  tally()

merge_data$Age_bucket <- as.factor(merge_data$Age_bucket)
```

```{r, message=FALSE, warning=FALSE}
merge_data[1:891,] %>%
  ggplot(aes(x = as.factor(Age_bucket), fill = as.factor(Survived)))+
  geom_bar()+
  facet_wrap(~Sex)+
  scale_fill_discrete(name = 'Survived')+
  xlab("Age Bucket")
```



Lets extract the first letter from the Cabin
```{r, message=FALSE, warning=FALSE}

length(unique(merge_data$Cabin))
merge_data$Cabin_Letter <- str_sub(merge_data$Cabin,1,1)

#creating a boolean variable

merge_data$Cabin_Available <- as.factor(!(merge_data$Cabin==""))

```


Fare
```{r, message=FALSE, warning=FALSE}
merge_data %>%
  ggplot(aes(x = Fare))+
  geom_histogram(fill = "#000099", color = "black")

sum(is.na(merge_data$Fare))

merge_data[which(is.na(merge_data$Fare)),]

#lets replace the fare by mean of Pclass = 3

merge_data[1044,]$Fare <- mean(subset(merge_data,Pclass==3)$Fare,na.rm=TRUE)

merge_data$Fare_bucket <- cut2(merge_data$Fare,g=5)
merge_data$Fare_bucket <- as.factor(merge_data$Fare_bucket)


merge_data[1:891,] %>%
  ggplot(aes(x = as.factor(Fare_bucket), fill = as.factor(Survived)))+
  geom_bar()+
  facet_wrap(~Sex)+
  scale_fill_discrete(name = 'Survived')+
  xlab("Fare Bucket")
```


Title
```{r, message=FALSE, warning=FALSE}

merge_data$Title <- merge_data$Title <- gsub('(.*, )|(\\..*)', '', merge_data$Name)
unique(merge_data$Title)

#We can see that there some titles that seem to have some luxurious power such as Sir, Major etc...


Power_Title <- c("Capt","Col","Don","Dona","Dr","Jonkheer","Lady","Major",
         "Mlle", "Mme","Rev","Sir","the Countess")

merge_data$Title[merge_data$Title %in% Power_Title] <- "Luxurious Title"
merge_data$Title <- as.factor(merge_data$Title)

```

Embarked
```{r, message=FALSE, warning=FALSE}

merge_data$Embarked %>% head()
merge_data$Embarked <- as.factor(merge_data$Embarked)
str(merge_data)
summary(merge_data$Embarked)
```


SibSp
```{r, message=FALSE, warning=FALSE}
merge_data$SibSp <- as.factor(merge_data$SibSp)
```

Ticket
```{r, message=FALSE, warning=FALSE}
library(plyr)
merge_data <- ddply(merge_data,.(Ticket),transform,Ticketsize=length(Ticket))
merge_data$Ticketsize <- as.factor(merge_data$Ticketsize)
merge_data <- merge_data  %>%
              arrange(PassengerId)

```



##### Data Modeling
```{r, message=FALSE, warning=FALSE}

#resplitting the merge data
train_merge <- merge_data[1:891,]
test_merge <- merge_data[892:1309,]

train_merge$Survived <- as.factor(train_merge$Survived)

#splitting the train merge dataset into a build and validate set
set.seed(123)
smp_size <- floor(0.75* nrow(train_merge))
train_ind <- sample(seq_len(nrow(train_merge)), size = smp_size)

train <- train_merge[train_ind, ]
test <- train_merge[-train_ind, ]


```


Logistic Regression
```{r, message=FALSE, warning=FALSE}

log.fit <- glm(Survived ~ Pclass + Sex + Age_bucket+ Fare_bucket +Cabin_Available + Title + Embarked + Ticketsize, family = binomial(link = logit), data = train)

summary(log.fit)


```

```{r, message=FALSE, warning=FALSE}
log_prediction <- predict(log.fit, newdata = test, type = "response")
log_prediction <- ifelse(log_prediction > 0.5 , 1 , 0)
logistic_classification_error <- mean(log_prediction != test$Survived)
logistic_accuracy <- 1 - logistic_classification_error

library(formattable)
print(paste0("The Accuracy for the Logistic Regression is: ", percent(logistic_accuracy)))

```

##### Let's use the Caret package to improvde our predictive abilities by using 10-fold cross validation

```{r, message=FALSE, warning=FALSE}
trainControl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
metric <- "Accuracy"
```


Logistic Regression
```{r, message=FALSE, warning=FALSE}
set.seed(123)
logistic.fit_2 <- train(Survived ~ Pclass + Sex + Age_bucket+ Fare_bucket +Cabin_Available + Title + Embarked + SibSp, data = train, method = "glm", metric = metric , trControl = trainControl)
```

KNN
```{r, message=FALSE, warning=FALSE}
library(caret)
set.seed(123)
knn.fit <- train(Survived ~ Pclass + Sex + Age_bucket+ Fare_bucket +Cabin_Available + Title + Embarked + SibSp, data = train, method = "knn", metric = metric , trControl = trainControl)

```

SVM
```{r}
set.seed(123)
svm.fit <- train(Survived ~ Pclass + Sex + Age_bucket+ Fare_bucket +Cabin_Available + Title + Embarked + SibSp, data = train, method = "svmRadial", metric = metric , trControl = trainControl)
```


Naives Bayes
```{r, message=FALSE, warning=FALSE}
set.seed(123)
naives.fit <- train(Survived ~ Pclass + Sex + Age_bucket+ Fare_bucket +Cabin_Available + Title + Embarked + SibSp, data = train, method = "nb", metric = metric , trControl = trainControl)

```

```{r, message=FALSE, warning=FALSE}
results <- resamples(list(Logistic_Regression =logistic.fit_2, KNN=knn.fit,
   NB=naives.fit, SVM=svm.fit))

summary(results)
```

```{r, message=FALSE, warning=FALSE}
dotplot(results)
```
From the plot above, we can see that SVm has a the highest accuracy

##### SVM Algorithm tuning

```{r, message=FALSE, warning=FALSE}
grid <- expand.grid(.sigma=c(0.005, 0.01, 0.015, 0.02), .C=seq(1, 10, by=1))

set.seed(1234)
svm.fit_2 <- train(Survived ~ Pclass + Sex + Age_bucket+ Fare_bucket + Cabin_Available + Title + Embarked + SibSp, data = train, method = "svmRadial", metric = metric , trControl = trainControl, tuneGrid = grid)

print(svm.fit_2)
```

```{r, message=FALSE, warning=FALSE}
plot(svm.fit_2)
```

##### Random Forest
```{r, message=FALSE, warning=FALSE}
rf.fit <- randomForest(factor(Survived) ~Pclass + Sex + Age_bucket + Fare_bucket +Cabin_Available + Title + Embarked + Ticketsize,
                              data = train ,nodesize=20)
print(rf.fit)
```
Based on the summary, the random forest model gives us an accuracy of 80.1%
```{r}
plot(rf.fit)
```


```{r, message=FALSE, warning=FALSE}
importance(rf.fit)
```

```{r}
rf.fit$confusion
```

Based on the random forest variable importance, lets only keep the following: Pclass + Sex + Age_bucket+ Fare_bucket Title + Ticketsize
```{r}
rf.fit2 <- randomForest(factor(Survived) ~Pclass + Sex + Fare_bucket + Cabin_Available + Ticketsize + Embarked + Title ,
                              data = train, nodesize=20)

rf.fit2$confusion
```


##### Conclusion: Logistic Regression is the winner 
```{r}
best_model <- rf.fit
prediction <- predict(best_model, test_merge)

submission <- data.frame(PassengerId=names(prediction),Survived=prediction)
submission %>% head()

write.csv(submission, file = "fourth_submission.csv", row.names = FALSE)

```



